### 2016-03-15 paper/icfp16-submission/evaluation.tex
misc edits everywhere
misc edits

We have implemented a prototype of our search procedure and trace
visualization for the pure subset of \ocaml, in a tool called \nanomaly.
In our implementation we instantiated \gensym\ with a simple random
generation of values, which we will show suffices for the
majority of type errors.

@@

\paragraph{Results}
\label{sec:results-complexity}
The results of the experiment are summarized in
Figure~\ref{fig:results-complexity}.
%
The average number of single-step reductions per trace is 31 for the
\ucsdbench\ dataset (35 for the \uwbench\ dataset) with a maximum of
2,745 (660 for \uwbench) and a median of 16 (also 16 for \uwbench).
%
The average number of jumps per trace is 7 (also 7 for \uwbench) with a
maximium of 353 (105 for \uwbench) and a median of 4 (also 4 for
\uwbench).
%
In both datasets at least 80\% of the traces contain at most 10 jumps.
%

@@

\paragraph{Infinite Types}
Our implementation does check for infinite types inside \forcesym, but
there are some degenerate cases where it is unable to detect
them. Consider, for example, the following buggy version of @replicate@.
% ,
% which is supposed to produce a list containing @n@ copies of @x@.
%
\begin{code}
  let rec replicate n x =
    if n <= 0 then
      []
    else
      replicate (n-1) [x]
\end{code}
%
This code produces a nested list (with @n@ levels of nesting) containing
a single copy of @x@, instead of a list with @n@ copies of @x@. \ocaml
detects a cyclic \hbox{@'a = 'a list@} constraint in the recursive call
and throws a type error, whereas \nanomaly happily % recurses @n@ times to
produces the nested list.  Strictly speaking, this function itself cannot
``go wrong'', the program would not get stuck until a \emph{client}
attempted to use the result expecting a flat list. But this is not very
satisfying as @replicate@ is clearly to blame. Furthermore, in our
experience, infinite-type errors are often difficult to %some of the more difficult ones to
debug (and to explain to novices), so better support for this scenario
would be useful.

\paragraph{Trace Explosion}
Though the average complexity of our generated traces is low in terms of
jumps, there are some extreme outliers. We cannot reasonably expect a
novice user to explore a trace containing 50+ terms and draw a
conclusion about which pieces contributed to the bug in their
program. Enhancing our visualization to slice out program paths relevant
to specific values, \eg\ in the style of~\cite{perera_functional_2012},
would likely help alleviate this issue, allowing users to
highlight a confusing value and ask: ``Where did this come from?''

>>>>

We have implemented a prototype of our search procedure and trace
visualization for the purely functional subset of \ocaml, in a tool called \nanomaly.
In our implementation we instantiated \gensym\ with a simple random
generation of values, which we will show suffices for the
majority of type errors.

@@

\paragraph{Results}
\label{sec:results-complexity}
The results of the experiment are summarized in
Figure~\ref{fig:results-complexity}.
%
The average number of single-step reductions per trace is 31 for the
\ucsdbench\ dataset (35 for the \uwbench\ dataset) with a maximum of
2,745 (660 for \uwbench) and a median of 16 (also 16 for \uwbench).
%
The average number of jumps per trace is 7 (also 7 for \uwbench) with a
maximium of 353 (105 for \uwbench) and a median of 4 (also 4 for
\uwbench).
%
In both datasets 80\% or more traces have at most 10 jumps.
%

@@

\paragraph{Infinite Types}
Our implementation does check for infinite types inside \forcesym, but
there are some degenerate cases where it is unable to detect
them. Consider, the following buggy @replicate@
%
\begin{code}
  let rec replicate n x =
    if n <= 0 then []
              else replicate (n-1) [x]
\end{code}
%
This code produces a nested list (with @n@ levels of nesting) containing
a single copy of @x@, instead of a list with @n@ copies of @x@. \ocaml
detects a cyclic \hbox{@'a = 'a list@} constraint in the recursive call
and throws a type error, whereas \nanomaly happily % recurses @n@ times to
produces the nested list.  Strictly speaking, this function itself cannot
``go wrong'', the program would not get stuck until a \emph{client}
attempted to use the result expecting a flat list. But this is not very
satisfying as @replicate@ is clearly to blame. Furthermore, in our
experience, infinite-type errors are often difficult to %some of the more difficult ones to
debug (and to explain to novices), so better support for this scenario
would be useful.

\paragraph{Trace Explosion}
Though the average complexity of our generated traces is low in terms of
jumps, there are some extreme outliers. We cannot reasonably expect a
novice user to explore a trace containing 50+ terms and draw a
conclusion about which pieces contributed to the bug in their
program. Enhancing our visualization to slice out program paths relevant
to specific values~\cite{perera_functional_2012},
would likely help alleviate this issue, allowing users to
highlight a confusing value and ask: ``Where did this come from?''

### 2016-03-15 paper/icfp16-submission/semantics.tex
misc edits

  % By Lemma~\ref{lem:fixme} we know that
  % $\vcompat{\resolve{\vhole{\thole}}{\vsu}}{\resolve{\thole}{\tsu}}$.
  %
  By Lemma~\ref{lem:vsu-ext} we know that
  $\tsu_{i} = \tsu_{i-1}[\thole_1 \mapsto t_1] \ldots [\thole_n \mapsto t_n]$.
  %
  We will assume, without loss of generality, that
  $\tsu_{i} = \tsu_{i-1}[\thole' \mapsto t']$.
  %
  Since $\tsu_{i-1}$ and $\tsu_{i}$ differ only in $\thole'$ but the resolved
  types differ,
  $\ptype{\trace_{i-1}}{f} = \intctx{\thole'}$
  and
  $\ptype{\trace_i}{f} = \intctx{t'}$.
  % $\resolve{\thole}{\tsu_{i-1}} = \intctx{\thole'}$
  % and
  % $\resolve{\thole}{\tsu_{i}} = \intctx{t'}$.
  %
  By Lemma~\ref{lem:context-compat},
  let $s$ be a concrete type such that $\intctx{s} = t$.
  %
  We show by case analysis on the evaluation rules that
  %
  $$\step{e_{i-1}}{\vsu_{i-1}}{\tsu_{i-1}[\thole' \mapsto s]}{\stuck}{\vsu}{\tsu}$$

>>>>

  % By Lemma~\ref{lem:fixme} we know that
  % $\vcompat{\resolve{\vhole{\thole}}{\vsu}}{\resolve{\thole}{\tsu}}$.
  %
  By Lemma~\ref{lem:vsu-ext} we know that
  $\tsu_{i} = \tsu_{i-1}[\thole_1 \mapsto t_1] \ldots [\thole_n \mapsto t_n]$.
  %
  We will assume, without loss of generality, that
  $\tsu_{i} = \tsu_{i-1}[\thole' \mapsto t']$.
  %
  Since $\tsu_{i-1}$ and $\tsu_{i}$ differ only in $\thole'$ but the resolved
  types differ, we have, for some type context $T$,
  $\ptype{\trace_{i-1}}{f} = \intctx{\thole'}$
  and
  $\ptype{\trace_i}{f} = \intctx{t'}$.
  %
  We prove, in Lemma~\ref{lem:context-compat},
  that \RJ{rephrase lamma here}.
  Thus, by Lemma~\ref{lem:context-compat}, let $s$ be a
  concrete type such that $\intctx{s} = t$.
  %
  We show by case analysis on the evaluation rules that
  %
  $$\step{e_{i-1}}{\vsu_{i-1}}{\tsu_{i-1}[\thole' \mapsto s]}{\stuck}{\vsu}{\tsu}$$
